Faster generic IND-CCA secure KEM using encrypt-then-MAC

### 2024-10-15 etm-transform/manuscript/main.tex
Cleaned up proof

\subsection{Contributions}
Our main contributions are threefold: \begin{itemize}
    \item \textbf{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} \textcolor{red}{\bf New IND-CCA secure KEM construction using encrypt-then-MAC.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC. In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC such as Poly1305 for further performance improvements.
    \item \textbf{$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} \textcolor{red}{\bf $\mlkemplus$: Efficient CCA-secure KEM based on ML-KEM.} We present $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines. Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
    \item \textbf{C implementation and performance comparison.}  \textcolor{red}{\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios. Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}). We also implemented the Kyber key exchange protocols \cite{DBLP:conf/eurosp/BosDKLLSSSS18} with $\mlkemplus$ and measured the round trip time (RTT) in realistic network settings. Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE). See Table \ref{tbl:rtt-summary} for a summary of round trip times.
\end{itemize}

\begin{table}[h]
    \centering
    % \footnotesize
    \small

    % TODO: Poly1305 may not be suitable for 192/256 bit security level
    % \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
    %     \hline
    %     & $\text{ML-KEM}^+$ \newline 512 
    %     & $\text{ML-KEM}^+$ \newline 768 
    %     & $\text{ML-KEM}^+$ \newline 1024 
    %     \\
    %     \hline
    %     Encap \newline (ccl/tick) 
    %     & 93157 \newline (+1.8\%) 
    %     & 146405 \newline (+7.3\%) 
    %     & 205763 \newline (+3.3\%) 
    %     \\
    %     \hline
    %     Decap \newline (ccl/tick) 
    %     & 33733 \newline (-72.2\%) 
    %     & 43315 \newline (-76.8\%) 
    %     & 51375 \newline (-79.1\%) 
    %     \\
    %     \hline
    %     CT size \newline (bytes) 
    %     & 784 \newline (+2.1\%) 
    %     & 1104 \newline (+1.5\%) 
    %     & 1584 \newline (+1.0\%) 
    %     \\
    %     \hline
    % \end{tabular}
    \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.4cm}|p{1.4cm}|}
       \hline
       & Encap \newline (ccl/tick) 
       & Decap \newline (ccl/tick) 
       & CT size \newline (bytes) \\
       \hline
       $\mlkemplus$-512 \newline (Poly1305)
       & 93157 \newline (+1.8\%)
       & 33733 \newline (-72.2\%)
       & 784 \newline (+2.1\%) \\
       \hline
       $\mlkemplus$-512 \newline (GMAC)
       & 97369 \newline (+6.5\%)
       & 37725 \newline (-68.9\%)
       & 784 \newline (+2.1\%) \\
       \hline
       $\mlkemplus$-512 \newline (CMAC)
       & 99739 \newline (+9.0\%)
       & 40117 \newline (-66.9\%)
       & 784 \newline (+2.1\%) \\
       \hline
       $\mlkemplus$-512 \newline (KMAC-256)
       & 101009 \newline (+10.4\%)
       & 40741 \newline (-66.4\%)
       & 784 \newline (+2.1\%) \\
       \hline
    \end{tabular}

@@

\begin{table}[h]
    \centering
    % \footnotesize
    \small

    % TODO: Poly1305 might not be suitable for 192/256 bit security
    % \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
    %     \hline
    %     &  $\text{ML-KEM}^+$ \newline 512 
    %     &  $\text{ML-KEM}^+$ \newline 768 
    %     &  $\text{ML-KEM}^+$ \newline 1024 
    %     \\
    %     \hline
    %     KE RTT \newline ($\us$) 
    %     &  70 \newline (-23.9\%) 
    %     &  99 \newline (-26.7\%) 
    %     &  138 \newline (-28.5\%) 
    %     \\
    %     \hline
    %     UAKE RTT \newline ($\us$) 
    %     &  103 \newline (-29.0\%) 
    %     &  144 \newline (-33.0\%) 
    %     &  202 \newline (-34.8\%) 
    %     \\
    %     \hline
    %     AKE RTT \newline ($\us$) 
    %     &  133 \newline (-39.5\%) 
    %     &  190 \newline (-35.4\%) 
    %     &  266 \newline (-48.0\%) 
    %     \\
    %     \hline
    % \end{tabular}
    \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.8cm}|p{1.8cm}|}
       \hline
       & KE RTT \newline ($\us$) 
       & UAKE RTT \newline ($\us$) 
       & AKE RTT \newline ($\us$) \\
       \hline
       $\mlkemplus$-512 \newline (Poly1305)
       & 70 \newline (-23.9\%)
       & 103 \newline (-29.0\%)
       & 133 \newline (-39.5\%) \\
       \hline
       $\mlkemplus$-512 \newline (GMAC)
       & 73 \newline (-20.7\%)
       & 106 \newline (-26.9\%)
       & 139 \newline (-36.8\%) \\
       \hline
       $\mlkemplus$-512 \newline (CMAC)
       & 75 \newline (-18.5\%)
       & 108 \newline (-25.5\%)
       & 143 \newline (-35.0\%) \\
       \hline
       $\mlkemplus$-512 \newline (KMAC-256)
       & 76 \newline (-17.4\%)
       & 109 \newline (-24.8\%)
       & 145 \newline (-34.1\%) \\
       \hline
    \end{tabular}

@@

    \caption{The $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
\end{figure}

@@

\section{The encrypt-then-MAC transformation}\label{sec:main-results}
In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM. Our scheme mainly differs from DHIES in its versatility and input requirement. Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:elgamal-is-ow-pca} that the Gap Diffie-Hellman assumption implies OW-PCA security). In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (M. Abdalla et al. originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme. The encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.

@@

In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC. In Section \ref{sec:elgamal-is-ow-pca}, we show that DHIES \cite{DBLP:conf/ctrsa/AbdallaBR01} is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.

@@

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow (\sk^\prime, z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $k \leftarrow G(\pk, m)$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow H(\pk, m, t)$
                \State $c \leftarrow (c^\prime, t)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $(\sk^\prime, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\pk, \hat{m})$
                \If{$\mac(\hat{k}, c^\prime) \neq t$}
                    \State $K \leftarrow H(z, t)$
                \Else
                    \State $K \leftarrow H(\pk, \hat{m}, t)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The encrypt-then-MAC transformation builds a KEM $\kem_\etm$ using a $\pke(\keygen, \encrypt, \decrypt)$, a $\mac$, and two hash functions $G, H$}\label{fig:etm-routines} 
\end{figure}

The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol. In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext $k \leftarrow G(m)$, then used to sign the unauthenticated ciphertext $c^\prime$. Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$ that would be transmitted to the peer. In the decapsulation routine, the decryption $\hat{m}$ of the unauthenticated ciphertext is used to re-derive the MAC key $\hat{k}$, which is then used to re-compute the tag $\hat{t}$. The ciphertext is considered valid if and only if the recomputed tag is identical to the input tag.

Here are some additional design rationales that have practical impact on security and performance but do not affect the security proof: \begin{itemize}
    \item \textbf{Removing de-randomization.} Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine. If the input PKE is already rigid, then the shared secret may be derived from hashing the PKE plaintext alone. However, if the input PKE is not rigid, then the shared secret must be derived from hashing both the PKE plaintext and the PKE ciphertext. 
    \item \textbf{Hashing the public key.} The public key goes into deriving both the MAC key and the shared secret for two main reasons. First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties. Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret. Second, if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext. Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption. Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
    \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext. However, we find it unnecessary to hash the entire ciphertext. Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements for shared secret derivation. On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
\end{itemize}

@@

However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed. This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see Figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.

\begin{figure}[H]
    \centering
    \begin{minipage}{0.45\textwidth}
    \begin{algorithm}[H]
        \caption*{$\pco(m, c)$}
        \begin{algorithmic}[1]
            \State $k \leftarrow G(m)$
            \State $t \leftarrow \mac(k, c)$
            \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \end{minipage}
    \caption{With the encrypt-then-MAC transformation, every chosen-ciphertext attack against the KEM implies a plaintext-checking attack against the underlying PKE}\label{fig:pco-using-decap-oracle}
\end{figure}

@@

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
We will prove Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game. A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}. From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$. A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.

>>>>

\subsection{Contributions}
Our main contributions are threefold: \begin{itemize}
    \item {\bf New IND-CCA secure KEM construction using encrypt-then-MAC.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC. In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC such as Poly1305 for further performance improvements.
    \item {\bf $\mlkemplus$: Efficient CCA-secure KEM based on ML-KEM.} We present $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines. Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
    \item {\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios. Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}). We also implemented the Kyber key exchange protocols \cite{DBLP:conf/eurosp/BosDKLLSSSS18} with $\mlkemplus$ and measured the round trip time (RTT) in realistic network settings. Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE). See Table \ref{tbl:rtt-summary} for a summary of round trip times.
\end{itemize}

\begin{table}[h]
    \centering
    \footnotesize
    % \small

    % TODO: Poly1305 may not be suitable for 192/256 bit security level
    \begin{tabular}{|p{1.2cm}|p{2.1cm}|p{2.6cm}|p{2.7cm}|}
        \hline
        & $\mlkemplus$-512 \newline (Poly1305)
        & $\mlkemplus$-768 \newline (KMAC 192-bit tag)
        & $\mlkemplus$-1024 \newline (KMAC 256-bit tag)
        \\
        \hline
        Encap \newline (ccl/tick) 
        & 91155 \newline (???) 
        & 150929 \newline (???) 
        & 228357 \newline (???) 
        \\
        \hline
        Decap \newline (ccl/tick) 
        & ??? \newline (???) 
        & ??? \newline (???) 
        & ??? \newline (???) 
        \\
        \hline
        CT size \newline (bytes) 
        & ??? \newline (???) 
        & ??? \newline (???) 
        & ??? \newline (???) 
        \\
        \hline
    \end{tabular}

    % Alternate version
    % \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.4cm}|p{1.4cm}|}
    %    \hline
    %    & Encap \newline (ccl/tick) 
    %    & Decap \newline (ccl/tick) 
    %    & CT size \newline (bytes) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (Poly1305)
    %    & 93157 \newline (+1.8\%)
    %    & 33733 \newline (-72.2\%)
    %    & 784 \newline (+2.1\%) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (GMAC)
    %    & 97369 \newline (+6.5\%)
    %    & 37725 \newline (-68.9\%)
    %    & 784 \newline (+2.1\%) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (CMAC)
    %    & 99739 \newline (+9.0\%)
    %    & 40117 \newline (-66.9\%)
    %    & 784 \newline (+2.1\%) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (KMAC-256)
    %    & 101009 \newline (+10.4\%)
    %    & 40741 \newline (-66.4\%)
    %    & 784 \newline (+2.1\%) \\
    %    \hline
    % \end{tabular}

@@

\begin{table}[h]
    \centering
    \footnotesize
    % \small

    % TODO: Poly1305 might not be suitable for 192/256 bit security
    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
        \hline
        &  $\text{ML-KEM}^+$ \newline 512 
        &  $\text{ML-KEM}^+$ \newline 768 
        &  $\text{ML-KEM}^+$ \newline 1024 
        \\
        \hline
        KE RTT \newline ($\us$) 
        &  ??? \newline (???) 
        &  ??? \newline (???) 
        &  ??? \newline (???) 
        \\
        \hline
        UAKE RTT \newline ($\us$) 
        &  ??? \newline (???) 
        &  ??? \newline (???) 
        &  ??? \newline (???) 
        \\
        \hline
        AKE RTT \newline ($\us$) 
        &  ??? \newline (???) 
        &  ??? \newline (???) 
        &  ??? \newline (???) 
        \\
        \hline
    \end{tabular}

    % Alternate version
    % \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.8cm}|p{1.8cm}|}
    %    \hline
    %    & KE RTT \newline ($\us$) 
    %    & UAKE RTT \newline ($\us$) 
    %    & AKE RTT \newline ($\us$) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (Poly1305)
    %    & 70 \newline (-23.9\%)
    %    & 103 \newline (-29.0\%)
    %    & 133 \newline (-39.5\%) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (GMAC)
    %    & 73 \newline (-20.7\%)
    %    & 106 \newline (-26.9\%)
    %    & 139 \newline (-36.8\%) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (CMAC)
    %    & 75 \newline (-18.5\%)
    %    & 108 \newline (-25.5\%)
    %    & 143 \newline (-35.0\%) \\
    %    \hline
    %    $\mlkemplus$-512 \newline (KMAC-256)
    %    & 76 \newline (-17.4\%)
    %    & 109 \newline (-24.8\%)
    %    & 145 \newline (-34.1\%) \\
    %    \hline
    % \end{tabular}

@@

    \caption{The $\kem^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
\end{figure}

@@

\section{The encrypt-then-MAC transformation}\label{sec:main-results}
In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM. Our scheme mainly differs from DHIES in its versatility and input requirement. Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:elgamal-is-ow-pca} that the Gap Diffie-Hellman assumption implies OW-PCA security). In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 for further performance improvements (M. Abdalla et al. originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}). The encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.

@@

In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC. In Section \ref{sec:elgamal-is-ow-pca}, we show that DHIES is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.

@@

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow (\sk^\prime, z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $k \leftarrow G(m)$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $c \leftarrow (c^\prime, t)$
                \State $K \leftarrow H(m, c)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $(\sk^\prime, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) = t$}
                    \State $K \leftarrow H(\hat{m}, c)$
                \Else
                    % This is directly copied from ML-KEM, DO NOT CHANGE
                    \State $K \leftarrow H(z, c)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The encrypt-then-MAC transformation builds a KEM $\kem_\etm$ using a $\pke(\keygen, \encrypt, \decrypt)$, a $\mac$, and two hash functions $G, H$}\label{fig:etm-routines} 
\end{figure}

The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol. In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext, then used to sign the unauthenticated ciphertext $c^\prime$. Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$ that would be transmitted to the peer. In the decapsulation routine, the decryption $\hat{m}$ of the unauthenticated ciphertext is used to re-derive the MAC key $\hat{k}$, which is then used to re-compute the tag $\hat{t}$. The ciphertext is considered valid if and only if the recomputed tag is identical to the input tag.

% Here are some additional design rationales that have practical impact on security and performance but do not affect the security proof: \begin{itemize}
%     \item \textbf{Removing de-randomization.} Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine. If the input PKE is already rigid, then the shared secret may be derived from hashing the PKE plaintext alone. However, if the input PKE is not rigid, then the shared secret must be derived from hashing both the PKE plaintext and the PKE ciphertext. 
%     \item \textbf{Hashing the public key.} The public key goes into deriving both the MAC key and the shared secret for two main reasons. First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties. Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret. Second, if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext. Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption. Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
%     \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext. However, we find it unnecessary to hash the entire ciphertext. Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements for shared secret derivation. On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
% \end{itemize}

@@

However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed. This means that a decapsulation oracle can be converted into a plaintext-checking oracle, so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.

@@

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
We will prove Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game. A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}. From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(\pk, m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, t)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$. A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{IND-CCA} game for $\kem_\etm$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \kem_\etm\texttt{.}\keygen()$
                \State $m^\ast \leftsample \mathcal{M}$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m^\ast)$
                \State $k^\ast \leftarrow G(m^\ast)$
                    \Comment{Game 0-1}
                \State $k^\ast \leftsample \mathcal{K}_\mac$
                    \Comment{Game 2-3}
                \State $t \leftarrow \mac(k^\ast, c^\prime)$
                \State $c^\ast \leftarrow (c^\prime, t)$
                \State $K_0 \leftarrow H(m^\ast, c^\ast)$
                    \Comment{Game 0-2}
                \State $K_0 \leftsample \mathcal{K}_\kem$
                    \Comment{Game 3}
                \State $K_1 \leftsample \mathcal{K}_\kem$
                \State $b \leftsample \{0,1\}$
                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(\pk, c^\ast, K_b)$
                    \Comment{Game 0}
                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(\pk, c^\ast, K_b)$
                    \Comment{Game 1-3}
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{Hash oracle $\mathcal{O}^G(m)$}
            \begin{algorithmic}[1]
                \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
                    \State \Return $\tilde{k}$
                \EndIf
                \State $k \leftsample \mathcal{K}_\mac$
                \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
                \State \Return $k$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{Decap oracle $\mathcal{O}^\decap(c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) = t$}
                    \State $K \leftarrow H(\hat{m}, c)$
                \Else 
                    \State $K \leftarrow H(z, c)$
                \EndIf 
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
        \vspace{-0.5cm}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}^\decap_1(c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
                    \tilde{m} = \decrypt(\sk^\prime, c^\prime)
                    \land \mac(\tilde{k}, c^\prime) = t
                $}
                    \State $K \leftarrow H(\tilde{m}, c)$
                \Else
                    \State $K \leftarrow H(z, c)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
        \vspace{-0.5cm}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}^H(m, c)$}
            \begin{algorithmic}[1]
                \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
                    \State \Return $\tilde{K}$
                \EndIf
                \State $K \leftsample \mathcal{K}_\kem$
                \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Sequence of games}\label{fig:etm-ind-cc2-sequence-of-games}
\end{figure}
